---
title: "Portfolio1"
author: "Nanna Høgsholt"
date: "2/3/2020"
output: html_document
---

Load libraries 
```{r}
library(pacman);p_load(tidyverse, ggplot2, lme4)
```

Load data into RStudioe
```{r}
sleepdata <- read.csv("sleepstudy.csv", header = T)
```

###Task 1: Plot the data: 
1.a: Get the data from one participant, e.g. using subset(). Make a linear regression for reaction time as a function of days of sleep deprivation, e.g. using lm(). Report the F-statistics.
```{r}
sleepdata_308 <- filter(sleepdata, Subject == 308)
summary(lm(Reaction ~ Days, data = sleepdata_308))
```

A linear regression reveals a significant effect of sleep deprivation on reaction times for Subject 308, F(1,8) = 17.12, p = 0.0033.

1.b: How many degrees of freedom does the relevant F-distribution have?

Degrees of freedom: 1 & 8 


1.c: At which F-value does a regression with this distribution become statistically significant (p<0.05)? 
```{r}
qf(0.95,1,8)
```
A regression with a similar distribution becomes statistically significant when F(1,8) = 5.32.  

1.d: Make a plot of the F-distribution.
```{r}
nn<-seq(0,20,len=100)
fdist<-df(nn,1,8)
plot(nn,fdist, type='l',xlab='F value - df(1,8)',ylab='Density')

cutoff = qf(0.95, 1,8)

#plotting a vertical line at the cutoff
lines(c(cutoff,cutoff),c(0,1),col='darkgreen')
#Add explanation for the line
text(cutoff+2,0.5,'p=0.05',col='darkgreen')
#draw F-value as point on the curve
points(17.12,df(17.12,1,8),col='red')
```

The plot visualizes one tail of the f distribution, the f-value at which the regression with df = 1 and 8 (red line) is significant and the relevant f-value (blue dot).

###Task 2. For all participant in the experiment: 
2.a: Find the coefficients (slope and intercept) for the regression for reaction time as a function of days of sleep deprivation (a hint for the solution: use group_by() in tidyverse or this function here:
https://stat.ethz.ch/R-manual/R-devel/library/nlme/html/lmList.html, hint2: pool=FALSE)
```{r}
sleepdata$Subject <- as.factor(sleepdata$Subject)
summary(lmList(Reaction~Days | Subject, data = sleepdata, pool=FALSE))


```



2.b: Combine both scatter plot and regression line in the same figure. You may also include all participants in one plot.

```{r}
ggplot(sleepdata, aes(x = sleepdata$Days, y = sleepdata$Reaction, color = Subject)) + geom_point() + geom_smooth(method = "lm", alpha = 0, fill = "Blue") + labs(x = "Days of sleep deprivation", y = "Reactiontime") + ggtitle("Plot 1")
```

```{r}
ggplot(sleepdata, aes(x = sleepdata$Days, y = sleepdata$Reaction, color = Subject)) + geom_point() + geom_smooth(method = "lm", alpha = 0.1, fill = "Blue") + labs(x = "Days of sleep deprivation", y = "Reactiontime") + facet_wrap(~Subject) + ggtitle("Plot 2")
```


2.c: Collect and report the inferential statistics for each participant in a table using t-statistics, including t-value, df and p-value.

```{r}
df <- data.frame(sum1$coefficients)
df %>% select(Estimate.Days, t.value.Days, Pr...t...Days)
```

From the table above, we see that only one participant (335) has a negative t-value, which was also evident in Plot 2, where the slope is negative. Moreover, it can be seen that the effect of sleep deprivation is diverse depending on the participant.

```{r}
sum1$df
```



2.d: How many individual participants display a statistically significant effect of sleep deprivation (p-values uncorrected for mulitple comparisons)?

From the inferential statistics in table in 2.c we see that 13 participants’ p < .05 meaning that more than half of the participants experienced a statistically significant development in reaction times as an effect of sleep deprivation.



####3. Across participants: 
3.a: Use the slopes you found for each participant in exercise 2 as a new dataset. Test the hypothesis that the slopes are larger than zero against the null-hypothesis that the slopes are zero (i.e. no differences in response time exist as a function of time).

```{r}
df_coef <- data.frame(sum1$coefficients)
t.test(df_coef$Estimate.Days)
```

A one sample t-test finds that the beta coefficients are significantly different from 0, t = 6.77, p < .001, M = 10.47. We therefore reject the null hypothesis that the slope is equal to zero, i.e. no development.

3.b: Justify your use of test statistics.

A one sample t-test allows us to compare a grand mean of the sample
with a set value, which in this case is 0 as we test the null hypothesis that the mean beta coefficient is different from zero.


3.c: Report inferential statistics.

A one sample t-test finds that the beta coefficients are significantly different
from 0, t = 6.77, p < .001, M = 10.47, 95% CI [7.21, 13.73], df = 17.

3.d: Make a plot with the mean reaction time and standard error bars for each day across participants and plot the averaged regression line in the same figure.

```{r}
ggplot(sleepdata, aes(x = Days, y = Reaction)) + geom_point(stat='summary', fun.y=mean) +
geom_errorbar(stat='summary', fun.data=mean_se, width = 0.7) + geom_smooth(method = "lm", alpha = 0, fill = "Blue") + ggtitle("Plot 3")
```

Plot 3 shows the general positive tendency of all 18 participants during the 10 days of the experiment. This suggests that sleep deprivation has an effect on mean reaction times.





